\papertitle

### 2014-03-20 final-report/paper.tex
programmer = developer

We propose \system, a visualization tool that helps database developers/users
explore and understand query execution and data movement in a distributed
database management system (DDBMS). \system provides quick insight into common
problems such as data skew or performance bottleneck by leveraging
visualization techniques to present: (1) data flow between query segments and
between workers, (2) query execution and operator dependencies, (3) cluster
utilization, (4) network utilization.

@@

Database query performance is often unpredictable, especially when the database is
distributed. This requires programmers or database developers to spend
significant amounts of time and effort isolating and debugging the root cause
of performance problems. These may include skewed data partitioning or run-time
issues, such as resource contention or adverse network conditions.

Our goal is to build a visualization tool designed to help a database developer understand the data flow and run-time
performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
database programmer or developer could quickly discover run-time bottlenecks as well as problematic data skew issues,
saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDBMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system is designed to help the Myria programmer answer the following questions regarding
their query execution. 
\system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
of query execution.

\begin{itemize}
\item \emph{How does the data flow through the system between operators?} 
The overall data flow in a DDBMS is represented by a \graph. It is important
for the programmer to understand the \graph in order to optimize it. To aid in this
understanding, we provide an interactive visualization of the \graph in \system.

@@

The first step in designing \system was to identify the possible causes that
could affect the execution performance of the query. After we identified
several such causes, that we will enumerate below, we leveraged visualization
techniques that we found fit to allow developers/programmers to get insight
into how the query was executed on the available resources and quickly
identify the source(s) of the performance penalty.

We found that the performance of the query might be affected by the following
factors, which greatly influenced the design of our visualization tool:
\begin{itemize}
   \item \emph{Wrong/unoptimized physical query plan.} We manually analyzed
several physical query plans and discovered that some queries were poorly
optimized. This led us to design a view that allows interactive exploration of
the physical query plan (mapped to a graph of query operators).
   \item \emph{Stragglers, tail latency, expensive query operators, poor
scheduling.} Our targeted systems are \emph{distributed} databases. Therefore,
as the query is executed in a distributed environment, the query coordinator
has to wait until all the workers finish executing their assigned tasks. To
give insight into how the tasks are distributed and executed at each workers,
we designed a view that shows the cluster utilization over time and at each
worker.
   \item \emph{Poor data partitioning, data skew.} Network traffic can cause
serious delay when executing a distributed query. Due to poor data
partitioning, workers might need to send unnecessary big amount of data between
each other. We designed a view that allows the developer/programmer to analyze
the data traffic generated while executing the query.
\end{itemize}

>>>>

We propose \system, a visualization tool that helps database developers
explore and understand query execution and data movement in a distributed
database management system (DDBMS). \system provides quick insight into common
problems such as data skew or performance bottleneck by leveraging
visualization techniques to present: (1) data flow between query segments and
between workers, (2) query execution and operator dependencies, (3) cluster
utilization, (4) network utilization.

@@

Database query performance is often unpredictable, especially when the database is
distributed. This requires database developers to spend
significant amounts of time and effort isolating and debugging the root cause
of performance problems. These may include skewed data partitioning or run-time
issues, such as resource contention or adverse network conditions.

Our goal is to build a visualization tool designed to help a database developer understand the data flow and run-time
performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
database developer could quickly discover run-time bottlenecks as well as problematic data skew issues,
saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDBMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system is designed to help the Myria developer answer the following questions regarding
their query execution. 
\system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system allows the user to visualize the following aspects
of query execution.

\begin{itemize}
\item \emph{How does the data flow through the system between operators?} 
The overall data flow in a DDBMS is represented by a \graph. It is important
for the developer to understand the \graph in order to optimize it. To aid in this
understanding, we provide an interactive visualization of the \graph in \system.

@@

The first step in designing \system was to identify the possible causes that
could affect the execution performance of the query. After we identified
several such causes, that we will enumerate below, we leveraged visualization
techniques that we found fit to allow developers to get insight
into how the query was executed on the available resources and quickly
identify the source(s) of the performance penalty.

We found that the performance of the query might be affected by the following
factors, which greatly influenced the design of our visualization tool:
\begin{itemize}
   \item \emph{Wrong/unoptimized physical query plan.} We manually analyzed
several physical query plans and discovered that some queries were poorly
optimized. This led us to design a view that allows interactive exploration of
the physical query plan (mapped to a graph of query operators).
   \item \emph{Stragglers, tail latency, expensive query operators, poor
scheduling.} Our targeted systems are \emph{distributed} databases. Therefore,
as the query is executed in a distributed environment, the query coordinator
has to wait until all the workers finish executing their assigned tasks. To
give insight into how the tasks are distributed and executed at each workers,
we designed a view that shows the cluster utilization over time and at each
worker.
   \item \emph{Poor data partitioning, data skew.} Network traffic can cause
serious delay when executing a distributed query. Due to poor data
partitioning, workers might need to send unnecessary big amount of data between
each other. We designed a view that allows the developer to analyze
the data traffic generated while executing the query.
\end{itemize}
